---
title: "Harris matrices using RPostgreSQL and DiagrammeR"
author: "Nick Gauthier"
date: "July 7, 2015"
output: 
  html_document: 
    highlight: haddock
    theme: cosmo
---

Here is an example workflow for taking stratigraphic information from the excavation database and generating on-the-fly Harris Matrices.  
  
We need three packages to accomplish this:  
1. RPostgreSQL for querying the DB and getting the results.  
2. dplyr for cleaning the query results from the DB so they can be read by the graphing function.  
3. DiagrammeR for generating Harris Matrices from the cleaned data.  

```{r warning = F, message = F}
library(RPostgreSQL)          # DB interface
library(dplyr)                # data cleaning
library(DiagrammeR)           # plotting
```
  
Connect to the excavation DB  

```{r}
con <- dbConnect("PostgreSQL",
                 host='192.168.2.100', 
                 port = '5432', 
                 dbname = 'gygaia', 
                 user = 'gygaiaro', 
                 password = 'gygaiaro')  
```
  
Choose an excavation area  

```{r}
area <- c('93', '545')
```
  
Generate a PostgreSQL query  

```{r}
query <- paste('SELECT excavation.contexts.subtable, excavation.contexts.context_number, excavation.contexts_stratigraphic_relationships.stratigraphic_relationship_type, excavation.contexts_stratigraphic_relationships.stratigraphic_relationship_subtype, excavation.contexts_stratigraphic_relationships.related_context_number FROM excavation.contexts INNER JOIN excavation.contexts_stratigraphic_relationships ON (excavation.contexts.context_number = excavation.contexts_stratigraphic_relationships.context_number) AND (excavation.contexts.area_northing = excavation.contexts_stratigraphic_relationships.area_northing) AND (excavation.contexts.area_easting = excavation.contexts_stratigraphic_relationships.area_easting) GROUP BY excavation.contexts.area_easting, excavation.contexts.area_northing, excavation.contexts.subtable, excavation.contexts.context_number, excavation.contexts_stratigraphic_relationships.stratigraphic_relationship_type, excavation.contexts_stratigraphic_relationships.stratigraphic_relationship_subtype, excavation.contexts_stratigraphic_relationships.related_context_number HAVING (((excavation.contexts.area_easting)=',area[1],') AND ((excavation.contexts.area_northing)=',area[2],')) ORDER BY excavation.contexts.context_number;', sep = '')
```
  
Query the database, make a data table

```{r, message = F, results = 'hide'}
db.out <- dbGetQuery(con, query) %>% tbl_df
dbDisconnect(con)        # close the DB connection
```

```{r echo = F, results = 'asis'}
knitr::kable(db.out[1:10,])
```
  
    
  
Clean the result so the plotting function can read it

```{r}
db.clean <- db.out %>% tbl_df %>%
     filter(stratigraphic_relationship_type == 'Earlier than') %>%
     select(context_number, related_context_number) %>%
     mutate_each(funs(as.character)) %>%
     transmute(relations = paste(context_number, related_context_number, sep = '-->')) %>%
     t %>% paste0(collapse = ';')
```

```{r echo = F, results ='asis'}
knitr::kable(db.clean)
```
  
    
Draw the Harris Matrix

```{r, fig.width = 9, fig.height = 4}
diagram.command <- paste("graph BT", db.clean, sep = ';') 
DiagrammeR(diagram.command)
```
  
   
   
   This is simple example output. The DiagrammeR package has a rich set of functions to map information from the DB onto the color, size, shape, etc of the outputs.